### 2020-06-14 manuscript.tex
Small comments

@@

Cryptic species are present throughout the tree of life. They are especially prevalent in ferns, because of processes such hybridization, polyploidy, and reticulate evolution. In addition, the simple morphology of ferns limits phenotypic variation and makes it difficult to detect cryptic species. The model fern genus \textit{Ceratopteris} has long been suspected to harbor cryptic diversity, specifically in the highly polymorphic \textit{C. thalictroides}. Yet no studies have included samples from throughout its pan-tropical range or utilized genomic sequencing, making it difficult to assess the full extent of cryptic variation within this genus. Here, we present the first multilocus phylogeny of the genus using reduced representation genomic sequencing (RADseq) and examine population structure, phylogenetic relationships, and ploidy level variation. We recover similar species relationships found in previous studies, find support for a named cryptic species as genetically distinct, and identify a novel genomic variation within two of the mostly broadly distributed species in the genus.

@@

\textbf{2.5. Population and genomic structure analysis}

Because species of \textit{Ceratopteris} are known to hybridize \autocite{hickok1974}, but also form reproductively isolated cryptic species \autocite{Masuyama2002}, we wanted to examine population structure of the genus across its distribution. The program {\small{STRUCTURE}} v. 2.3.4 \autocite{Pritchard2000} is designed to determine admixture (hybridization) between populations of one or several closely related taxa. The program assumes that each individual's genome is a mosaic from \textit{K} source populations and uses genotype assignments from SNPs to infer population structure and admixture. We utilized a set of 8478 SNPs, created in \textit{pyrad} by selecting one SNP per loci. We ran {\small{STRUCTURE}} for \textit{K} = 2 - 6 with 50 chains for each \textit{K}. We then used {\small{CLUMPAK}} \autocite{Kopelman2015} to process the {\small{STRUCTURE}} output and estimate the best \textit{K} values \autocite{Evanno2005, Pritchard2000}.

@@

We ran {\small{TETRAD}} via Python 2.7.16 (www.python.org) for 50 samples of \textit{Ceratopteris}, and one sample of \textit{Acrostichum aureum} as the outgroup. Our input consisted of 51 individuals and 26891 SNPs; we ran 100 bootstrap iterations for the final consensus tree. As a comparison, we also ran {\small{RAxML}} v. 8.2.11 \autocite{Stamatakis2014} with 1000 bootstrap iterations. We then plotted the final trees in R using the package phytools \autocite{Revell2012} and custom plotting functions.

@@

Unfortunately, all samples of \textit{Ceratopteris richardii} (including \textit{Hnn}) either failed in sequencing, or had very poor sequencing quality. After an initial examination of the results of these samples, we chose to remove them from our final analyses. We did not want to attempt to make any inferences about this important species, nor the lab strain, with poor-quality data.

@@

 The {\small{RAxML}} tree shows a very similar structure to the {\small{TETRAD}} tree. The only major difference was that instead of New World \textit{C. thalictroides} and \textit{C. gaudichaudii} forming a monophyletic group, the former comes out as sister to all other species except \textit{C. pteridoides}. Graphical comparison of the two trees is included in Supplementary data 2.

@@

We attempted to create a pseudo-individual to include in our analysis using the \textit{Ceratopteris richardii} genome assembly. We were able to obtain 9,047,554 simulated raw sequences utilizing an \textit{in silico} RADseq digest. These reads, in combination with the rest of our \textit{in vivo} data, were processed with the \textit{ipyrad} pipeline. The \textit{ipyrad} pipeline consists of 7 steps, each performing one aspect of RADseq data processing \autocite{Eaton2020}. Our pseudo-individual of \textit{C. richardii} performed nearly identically to all of the other samples until Step 6. Here, similar sequences are clustered across samples, and then these shared sequences are mapped to a reference (our reference was a \textit{de novo} assembly, because the quality and coverage of the \textit{C. richardii} genome was too poor to serve as a reference). This requires that all individuals are related relatively close to one another, so that they share a large number of sequences. Very few sequences were retained for our pseudo-individual of \textit{C. richardii}, indicating that the sequences obtained are too different from the \textit{in vivo} samples. In order to retain enough SNPs to run STRUCTURE, we required that only 3 individuals needed to have data at a given locus for it to be included in the final data set. In our entirely \textit{in vivo} dataset, this minimum number of individuals was 30; we designated such a high number to reduce missing data. In our dataset with the pseudo-individual of \textit{C. richardii}, a vast majority of our retained loci had missing data.

The differences between the pseudo-individual derived from the \textit{Ceratopteris richardii} and our other samples is mostly likely due to the fact that the current assembly (v. 1.1) only represents about one third of the total genome, and consists of mostly short reads. Long reads that can span the many repetitive elements in the genome comprise only 0.03\% of the final assembly \autocite{Marchant2019}. Because of these factors, the current genome assembly contains a small amount of short sequences that are difficult to accurately compare to the fragments generated by \textit{in vivo} RADseq. We have included the results from our \textit{in silico} RADseq digest in Supplementary data 2, but do not present them in the main manuscript.

@@

The authors thank the University of Wisconsin Biotechnology Center DNA Sequencing Facility for providing DNA extraction, library prep, and DNA sequencing facilities and services. We would also like to thank the University of Utah Center for High-Performance Computing, particularly Anita Orendt, for providing computational resources for data analyses. Thank you to Ashley Field and Tzu-Tong Kao helping with field work; Jacob S. Suissa and David Barrington for sending herbarium specimens; and Ryan Choi and Jacob S. Suissa for their thoughtful comments and edits to the manuscript. SPK is funded by a National Science Foundation Graduate Research Fellowship and the Joseph E. Greaves Endowed Scholarship from Utah State University. WDP and the Pearse lab are funded by National Science Foundation ABI‐1759965, NSF EF‐1802605 and United States Department of Agriculture Forest Service agreement 18‐CS‐11046000‐041.

>>>>

% An example of how to make your life easier for odd formatting
\usepackage{xspace}
\newcommand{\structure}{\small{STRUCTURE}\xspace}

@@

Cryptic species are present throughout the tree of life. They are especially prevalent in ferns, because of processes such hybridization, polyploidy, and reticulate evolution. In addition, the simple morphology of ferns limits phenotypic variation and makes it difficult to detect cryptic species. The model fern genus \textit{Ceratopteris} has long been suspected to harbor cryptic diversity, specifically in the highly polymorphic \textit{C. thalictroides}. Yet no studies have included samples from throughout its pan-tropical range or utilized genomic sequencing, making it difficult to assess the full extent of cryptic variation within this genus. Here, we present the first multilocus phylogeny of the genus using reduced representation genomic sequencing (RADseq), and examine population structure, phylogenetic relationships, and ploidy level variation. We recover similar species relationships found in previous studies, find support for a named cryptic species as genetically distinct
% WDP: "named cryptic species" is a bit ambiguous - name it?...
, and identify a novel genomic variation within two of the mostly broadly distributed species in the genus.

@@

\textbf{2.5. Population and genomic structure analysis}
% WDP: See top of MS for how \structure works; I sense you might like this trick
Because species of \textit{Ceratopteris} are known to hybridize \autocite{hickok1974}, but also form reproductively isolated cryptic species \autocite{Masuyama2002}, we wanted to examine population structure of the genus across its distribution. The program \structure v. 2.3.4 \autocite{Pritchard2000} is designed to determine admixture (hybridization) between populations of one or several closely related taxa. The program assumes that each individual's genome is a mosaic from \textit{K} source populations and uses genotype assignments from SNPs to infer population structure and admixture. We utilized a set of 8478 SNPs, created in \textit{pyrad} by selecting one SNP per loci. We ran \structure for \textit{K} = 2 - 6 with 50 chains for each \textit{K}. We then used {\small{CLUMPAK}} \autocite{Kopelman2015} to process the \structure output and estimate the best \textit{K} values \autocite{Evanno2005, Pritchard2000}.

@@

We ran {\small{TETRAD}} via Python 2.7.16 (www.python.org) for 50 samples of \textit{Ceratopteris}, and one sample of \textit{Acrostichum aureum} as the outgroup. Our input consisted of 51 individuals and 26891 SNPs; we ran 100 bootstrap iterations for the final consensus tree. As a comparison, we also ran {\small{RAxML}} v. 8.2.11 \autocite{Stamatakis2014} with 1000 bootstrap iterations.
% WDP: What was the exact call? 1000 rapid bootstrap (which means, I think, 200 additional full ML searches?) searched? 1000 independent, full ML searches?
We then plotted the final trees in R using the package phytools \autocite{Revell2012} and custom plotting functions.

@@

Unfortunately, all samples of \textit{Ceratopteris richardii} (including \textit{Hnn}) either failed in sequencing, or had very poor sequencing quality. After an initial examination of the results of these samples, we chose to remove them from our final analyses. We did not want to attempt to make any inferences about this important species, nor the lab strain, with poor-quality data.
% WDP: Why are you making no reference to the supplementary materials here? Your results were robust either way - mention that here.

@@

 The {\small{RAxML}} tree shows a very similar structure to the {\small{TETRAD}} tree. The only major difference was that instead of New World \textit{C. thalictroides} and \textit{C. gaudichaudii} forming a monophyletic group, the former comes out as sister to all other species except \textit{C. pteridoides}. Graphical comparison of the two trees is included in Supplementary data 2.
 % WDP: Give a justification for these being in the supplementary materials; the usual one is "since these results are broadly similar we present them in the supplementary materials" or "since these results are qualitatively identical" or "do not affect conclusions" or some-such nonsense.

@@

We attempted to create a pseudo-individual to include in our analysis using the \textit{Ceratopteris richardii} genome assembly, and we were able to obtain 9,047,554 simulated raw sequences from our \textit{in silico} RADseq digest. These reads, in combination with the rest of our \textit{in vivo} data, were processed with the \textit{ipyrad} pipeline. The \textit{ipyrad} pipeline consists of 7 steps, each performing one aspect of RADseq data processing \autocite{Eaton2020}. Our pseudo-individual of \textit{C. richardii} performed nearly identically to all of the other samples until Step 6. Here, similar sequences are clustered across samples, and then these shared sequences are mapped to a reference (our reference was a \textit{de novo} assembly, because the quality and coverage of the \textit{C. richardii} genome was too poor to serve as a reference).
% WDP: Condense the above; this is too much background (and some of this is methods, I sense, so if you really can't cut it move it to the methods).
This requires that all individuals are related relatively close to one another, so that they share a large number of sequences. Very few sequences were retained for our pseudo-individual of \textit{C. richardii}, indicating that the sequences obtained are too different from the \textit{in vivo} samples. In order to retain enough SNPs to run STRUCTURE, we required that only 3 individuals needed to have data at a given locus for it to be included in the final data set. In our entirely \textit{in vivo} dataset, this minimum number of individuals was 30; we designated such a high number to reduce missing data. In our dataset with the pseudo-individual of \textit{C. richardii}, a vast majority of our retained loci had missing data.

The differences between the pseudo-individual derived from the \textit{Ceratopteris richardii} and our other samples is mostly likely due to the fact that the current assembly (v. 1.1) only represents about one third of the total genome, and consists of mostly short reads. Long reads that can span the many repetitive elements in the genome comprise only 0.03\% of the final assembly \autocite{Marchant2019}. Because of these factors, the current genome assembly contains a small amount of short sequences that are difficult to accurately compare to the fragments generated by \textit{in vivo} RADseq. We have included the results from our \textit{in silico} RADseq digest in Supplementary data 2, but do not present them in the main manuscript.
% WDP: SAY WHAT THE RESULTS SHOW - they don't change your conclusions, so make that clear. You're making it sound like you're hiding something when, in fact, running all of this through makes your conclusions even clearer.

@@

The authors thank the University of Wisconsin Biotechnology Center DNA Sequencing Facility for providing DNA extraction, library prep, and DNA sequencing facilities and services. We would also like to thank the University of Utah Center for High-Performance Computing, particularly Anita Orendt, for providing computational resources for data analyses. Thank you to Ashley Field and Tzu-Tong Kao helping with field work; Jacob S. Suissa and David Barrington for sending herbarium specimens; and Ryan Choi and Jacob S. Suissa for their thoughtful comments and edits to the manuscript. SPK is funded by a National Science Foundation Graduate Research Fellowship and the Joseph E. Greaves Endowed Scholarship from Utah State University. WDP and the Pearse lab are funded by National Science Foundation ABI‐1759965, NSF EF‐1802605, United States Department of Agriculture Forest Service agreement 18‐CS‐11046000‐041, and UK NERC grant NE/V009710/1.
